Yeast RNA-seq 48 sample simulation study

### 2018-07-24 datasets/RNA-Seq/polyester-simulation.Rmd
Weight DE sampling by informative covariate; switch to mclapply

# # set up parallel backend:
library(BiocParallel)
nCores <- 20
register(MulticoreParam(workers = nCores))
```

@@

```{r}
#' @param sampleSize is the number of samples in each condition
#' @param nDE is the number of DE genes
#' @param bd is the bench design object
#' @param uninformativeCovariate logical indicating whether to replace overall mean 
#'  independent covariate with a randomly generated (uninformative) covariate. 
#'  Default is FALSE.
#' @param rseed random seed
simulateOneSplit <- function(X, rseed, 
                             nDE, sampleSize, bd, uninformativeCovariate = FALSE){
  ngenes <- 10000

  uninf_covar <- rnorm(ngenes)

@@

  gs <- sample(1:nrow(counts), ngenes, replace = TRUE)
  mu <- paramPolyester$mu[gs]
  p0 <- paramPolyester$p0[gs]

@@

  # DE simulation
  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    DE <- sample(1:nrow(dat_alt), nDE)
    truth[DE] <- TRUE

@@

if (!file.exists(resfile_n5)){
  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
                    nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null5, file=resfile_n5)
} else {
  null5 <- readRDS(file=resfile_n5)
}
```

@@

if (!file.exists(resfile_n10)){
  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit,
                     nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null10, file=resfile_n10)
} else {
  null10 <- readRDS(file=resfile_n10)
}
```

@@

if (!file.exists(resfile_d5)){
  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de5, file=resfile_d5)
} else {
  de5 <- readRDS(file=resfile_d5)
}
```

@@

if (!file.exists(resfile_d10)){
  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de10, file=resfile_d10)
} else {
  de10 <- readRDS(file=resfile_d10)
}

@@

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                    nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(null5, file=resfile_n5_uninfCov)
} else {
  null5 <- readRDS(file=resfile_n5_uninfCov)
}
```

@@

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                     nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(null10, file=resfile_n10_uninfCov)
} else {
  null10 <- readRDS(file=resfile_n10_uninfCov)
}
```

@@

if (!file.exists(resfile_d5_uninfCov)){
  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, 
                  bd=bd, uninformativeCovariate = TRUE)
  saveRDS(de5, file=resfile_d5_uninfCov)
} else {
  de5 <- readRDS(file=resfile_d5_uninfCov)
}
```

@@

if (!file.exists(resfile_d10_uninfCov)){
  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(de10, file=resfile_d10_uninfCov)
} else {
  de10 <- readRDS(file=resfile_d10_uninfCov)
}

>>>>

# # set up parallel backend:
library(parallel)
nCores <- 20
```

@@

```{r}
#' @param sampleSize is the number of samples in each condition
#' @param nDE is the number of DE genes
#' @param bd is the bench design object
#' @param uninformativeCovariate logical indicating whether to replace overall mean 
#'  independent covariate with a randomly generated (uninformative) covariate. 
#'  Default is FALSE.
#' @param rseed random seed
#' @param ngenes the number of genes to simulate. If this number is greater than 
#'  the number of genes in the count matrix, then the polyester parameters will be
#'  drawn with replacement from the available pool in order to obtain the requested
#'  number of genes.
simulateOneSplit <- function(X, rseed, 
                             nDE, sampleSize, bd, uninformativeCovariate = FALSE,
                             ngenes = nrow(counts)){
  uninf_covar <- rnorm(ngenes)

@@

  gs <- sample(1:nrow(counts), ngenes, replace = as.logical(ngenes > nrow(counts)))
  mu <- paramPolyester$mu[gs]
  p0 <- paramPolyester$p0[gs]

@@

  # DE simulation
  if(nDE > 0){
    wts <- rep(1, nrow(dds_test))
    if(!uninformativeCovariate)
      wts <- log10(rowMeans(counts)+1)

    DE <- sample(1:nrow(dat_alt), nDE, prob = wts)
    truth[DE] <- TRUE

@@

if (!file.exists(resfile_n5)){
  null5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
  saveRDS(null5, file=resfile_n5)
} else {
  null5 <- readRDS(file=resfile_n5)
}
```

@@

if (!file.exists(resfile_n10)){
  null10 <- mclapply(X=1:B, rseed=837, FUN=simulateOneSplit,
                     nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(null10, file=resfile_n10)
} else {
  null10 <- readRDS(file=resfile_n10)
}
```

@@

if (!file.exists(resfile_d5)){
  de5 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(de5, file=resfile_d5)
} else {
  de5 <- readRDS(file=resfile_d5)
}
```

@@

if (!file.exists(resfile_d10)){
  de10 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
  saveRDS(de10, file=resfile_d10)
} else {
  de10 <- readRDS(file=resfile_d10)
}

@@

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                    nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(null5, file=resfile_n5_uninfCov)
} else {
  null5 <- readRDS(file=resfile_n5_uninfCov)
}
```

@@

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                     nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(null10, file=resfile_n10_uninfCov)
} else {
  null10 <- readRDS(file=resfile_n10_uninfCov)
}
```

@@

if (!file.exists(resfile_d5_uninfCov)){
  de5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, 
                  bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(de5, file=resfile_d5_uninfCov)
} else {
  de5 <- readRDS(file=resfile_d5_uninfCov)
}
```

@@

if (!file.exists(resfile_d10_uninfCov)){
  de10 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE, mc.cores = nCores)
  saveRDS(de10, file=resfile_d10_uninfCov)
} else {
  de10 <- readRDS(file=resfile_d10_uninfCov)
}

### 2018-07-24 datasets/RNA-Seq/yeast-simulation.Rmd
Weight DE sampling by informative covariate; switch to mclapply

# set up parallel backend
library(BiocParallel)
nCores <- 20
register(MulticoreParam(workers = nCores))
```

@@

  # select a random subset of 20 WT samples
  dds_test <- dds[,colData(dds)$condition == "Snf2"]
  dds_test <- dds_test[,sample(1:ncol(dds_test), sampleSize*2)]

@@

  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    DE <- sample(1:nrow(dds_test), nDE)
    truth[DE] <- TRUE

    # randomly sample a log2FC from original FCs (without regard to DE)
    counts_new <- counts(dds_test)
    log2FC <- rep(0, nrow(dds_test))
    log2FC[DE] <- resFULL$log2FoldChange[DE]

@@

  dds_test <- DESeq(dds_test, parallel = FALSE)
  resultsNames(dds_test) # lists the coefficients
  resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = FALSE)

@@

```{r}
B <- 100
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

@@

if (!file.exists(resfile_n5)){
  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null5, file = resfile_n5)
}else{
  null5 <- readRDS(file = resfile_n5)
}
```

@@

if (!file.exists(resfile_n10)){
  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null10, file = resfile_n10)
}else{
  null10 <- readRDS(file = resfile_n10)
}

@@

if (!file.exists(resfile_d5)){
  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de5, file = resfile_d5)
}else{
  de5 <- readRDS(file = resfile_d5)
}

# pvalue histograms
plotfile <- file.path(datdir, "de5_pvalhists.pdf")
if (!file.exists(plotfile)){
  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
                   pvalHists = TRUE)
  pdf(plotfile, width=8, height=4)
  for(i in 1:length(hists)){
    print(hists[[i]])
  }
  dev.off()
}
```

@@

if(!file.exists(resfile_d10)){
  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de10, file = resfile_d10)
}else{
  de10 <- readRDS(file = resfile_d10)
}

# pvalue histograms
plotfile <- file.path(datdir, "de10_pvalhists.pdf")
if (!file.exists(plotfile)){
  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
                   pvalHists = TRUE)
  pdf(plotfile, width=8, height=4)
  for(i in 1:length(hists)){
    print(hists[[i]])
  }
  dev.off()
}
```

@@

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE)
  saveRDS(null5, file = resfile_n5_uninfCov)
}else{
  null5 <- readRDS(file = resfile_n5_uninfCov)
}
```

@@

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE)
  saveRDS(null10, file = resfile_n10_uninfCov)
}else{
  null10 <- readRDS(file = resfile_n10_uninfCov)
}
```

@@

if (!file.exists(resfile_d5_uninfCov)){
  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE)
  saveRDS(de5, file = resfile_d5_uninfCov)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov)
}
```

@@

if(!file.exists(resfile_d10_uninfCov)){
  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE)
  saveRDS(de10, file = resfile_d10_uninfCov)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov)
}
```

>>>>

# set up parallel backend
library(parallel)
nCores <- 20
```

@@

  # select a random subset of 20 WT samples
  dds_test <- dds[,sample(1:ncol(dds), sampleSize*2)]

@@

  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    wts <- rep(1, nrow(dds_test))
    if(!uninformativeCovariate)
      wts <- log10(rowMeans(counts(dds_test, normalize = TRUE))+1)
    DE <- sample(1:nrow(dds_test), nDE, prob = wts)
    truth[DE] <- TRUE

    # randomly sample a log2FC from original FCs (without regard to DE)
    counts_new <- counts(dds_test)
    log2FC <- rep(0, nrow(dds_test))
    log2FC[DE] <- resFULL$log2FoldChange[pzero < 0.5][DE]

@@

  dds_test <- DESeq(dds_test, parallel = FALSE)
  resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = FALSE)

@@

```{r}
B <- 100
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05

# only keep one condition for subsetting in simulations that follow 
dds <- dds[,colData(dds)$condition == "Snf2"]
```

@@

if (!file.exists(resfile_n5)){
  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(null5, file = resfile_n5)
}else{
  null5 <- readRDS(file = resfile_n5)
}
```

@@

if (!file.exists(resfile_n10)){
  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(null10, file = resfile_n10)
}else{
  null10 <- readRDS(file = resfile_n10)
}

@@

if (!file.exists(resfile_d5)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(de5, file = resfile_d5)
}else{
  de5 <- readRDS(file = resfile_d5)
}

# pvalue histograms
plotfile <- file.path(datdir, "de5_pvalhists.pdf")
if (!file.exists(plotfile)){
  hists <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
                   pvalHists = TRUE, mc.cores = nCores)
  pdf(plotfile, width=8, height=4)
  for(i in 1:length(hists)){
    print(hists[[i]])
  }
  dev.off()
}
```

@@

if(!file.exists(resfile_d10)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
  saveRDS(de10, file = resfile_d10)
}else{
  de10 <- readRDS(file = resfile_d10)
}

# pvalue histograms
plotfile <- file.path(datdir, "de10_pvalhists.pdf")
if (!file.exists(plotfile)){
  hists <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
                   pvalHists = TRUE, mc.cores = nCores)
  pdf(plotfile, width=8, height=4)
  for(i in 1:length(hists)){
    print(hists[[i]])
  }
  dev.off()
}
```

@@

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, mc.cores = nCores)
  saveRDS(null5, file = resfile_n5_uninfCov)
}else{
  null5 <- readRDS(file = resfile_n5_uninfCov)
}
```

@@

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(null10, file = resfile_n10_uninfCov)
}else{
  null10 <- readRDS(file = resfile_n10_uninfCov)
}
```

@@

if (!file.exists(resfile_d5_uninfCov)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(de5, file = resfile_d5_uninfCov)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov)
}
```

@@

if(!file.exists(resfile_d10_uninfCov)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(de10, file = resfile_d10_uninfCov)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov)
}
```
